Luminance constancy under fixed geometry

### 2018-04-26 Draft4/draft4.tex
edits and figures

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of LRV under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived luminance of an object. We have developed a software that generates naturalistic multispectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the luminance of a specific object in the scene. Next, we simulate the response of the retinal cones to these multispectral images using an accurate model of the early visual system. We then use supervised learning methods on these labeled cone responses to identify the computations that lead to accurate luminance estimation. We show that if only either target, or background, or illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the luminance within about 15\% RMSE.}

@@

%% Fix the title capitalization in a and b
\subsubsection{Illumination spectra}
% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Illumination spectra:} (a) Granada dataset. The spectra are normalized by the mean over wavelength. (b) A representative sample of spectra generated using normalized Granada dataset. (c) CIE xy chromaticity diagram of the normalized dataset and the random samples. (d) sRGB renditions of the samples from Figure~\ref{fig:illuminantSamples}.}
\label{fig:illuminant}
\end{figure}

@@

Because our multivariate Gaussian model is based on the rescaled spectra, it does not embody the large variation in overall intensity of natural daylights.
To obtain illuminant spectra with intensity variation similar to the Granada dataset, we scaled each randomly generated relative spectrum.
We obtained each scale factor by randomly sampling (with replacement) a spectrum from the Granda dataset and taking its mean. 
Color swatches of samples of illuminant spectra are shown in Figure~\ref{fig:sRGBIlluminant}.

\subsubsection{Surface reflectance spectra}
%% Fix the title capitalization in a and b
% Figure 7 Surface Reflectance Data and Model
\begin{figure}
\centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
            \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_a.pdf}
        \label{fig:reflectanceSpectra}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_b.pdf}
        \label{fig:reflectanceSamples}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
    \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_c.pdf}
        \label{fig:xyChroReflectance}
    \end{subfigure}    
    \centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_d.pdf}
        \label{fig:backgroundSwatches}
    \end{subfigure}
    \caption{{\bf Surface reflectance spectra:} (a) Surface reflectance spectra from the Munsell and Vrhel datasets.(b) Representative samples of object surface reflectances. (c) CIE xy chromaticity diagram of the Munsell and Vrhel datasets (black) and samples of surface reflectances (red). (d) sRGB renditions of the samples of object surface reflectance under CIE D65 daylight.}
\label{fig:surfaceReflectanceGeneration}
\end{figure}

@@

\begin{figure}
\centering
            \begin{subfigure}[b]{0.25 \textwidth}
        \caption{LRV estimates}
        \includegraphics[width=\textwidth, trim={0 0 0 1.3cm},clip]{../FiguresDraft4/Figure10/Figure10_a.pdf}
        \label{fig:case1Estimates}
    \end{subfigure} 
        \begin{subfigure}[b]{0.26 \textwidth}
        \caption{Receptive field response}
        \includegraphics[width=\textwidth, trim={0 3mm 0 15mm},clip]{../FiguresDraft4/Figure10/Figure10_d.pdf}
        \label{fig:case1RFResponse}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.28 \textwidth}
	\caption{AMA receptive fields}
	\includegraphics[width=1.07\textwidth, trim={0 -0.6cm 0 -0.3cm}]{../FiguresDraft4/Figure10/Figure10_c.pdf}
	\label{fig:case1RFs}
    \end{subfigure}   
    \caption{{\bf RF response and luminance estimates for condition 1:} (a) Photoreceptor response projected along the first two AMA receptive fields for images of condition 1. We show the response for the images in the training set. Each response cloud represents the response of images corresponding to one LRV. The response clouds are approximated by a multivariate Gaussian whose mean and variance are approximated by the ellipses shown in the figure. (b)  Estimated v/s assigned target LRV for the image in condition 1. Solid lines show the mean estimate, the filled region in light color shows $\pm$ 1 standard deviation. Relative error of estimation is 0.8\% (Linear regression) and 1.1\% (AMA) . The diagonal unit slope (blue dotted) line is for reference. The standard deviations are too small to be visible.}
\label{fig:Condition1}
\end{figure}

@@

%% Figure 12: Results for Condition 3
\begin{figure}
\centering
            \begin{subfigure}[b]{0.25 \textwidth}
        \caption{LRV estimates}
        \includegraphics[width=\textwidth, trim={0 0 0 1.3cm},clip]{../FiguresDraft4/Figure12/Figure12_a.pdf}
        \label{fig:case3Estimates}
    \end{subfigure} 
        \begin{subfigure}[b]{0.26\textwidth}
        \caption{Receptive field response}
        \includegraphics[width=\textwidth, trim={0 3mm 0 15mm},clip]{../FiguresDraft4/Figure12/Figure12_d.pdf}
        \label{fig:case3RFResponse}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.28 \textwidth}
	\caption{AMA receptive fields}
	\includegraphics[width=1.07\textwidth, trim={0 -0.6cm 0 -0.5cm}]{../FiguresDraft4/Figure12/Figure12_c.pdf}
	\label{fig:case3RFs}
    \end{subfigure}       \caption{{\bf RF response and luminance estimates for condition 3:} (a) Photoreceptor response projected along the first two AMA receptive fields for images of condition 1. We show the response for the images in the training set. Each response cloud represents the response of images corresponding to one LRV. The response clouds are approximated by a multivariate Gaussian whose mean and variance are approximated by the ellipses shown in the figure. (b)  Estimated v/s assigned target LRV for the image in condition 1. Solid lines show the mean estimate, the filled region in light color shows $\pm$ 1 standard deviation. The relative error of estimation is 22.9\%  for linear regression and 13.0\% for AMA. The diagonal unit slope (blue dotted) line is for reference. The standard deviations are too small to be visible.}
\label{fig:Condition3}
\end{figure}

@@

Figure~\ref{fig:barGraphs} summarizes luminance constancy performance across the conditions, for both AMA and linear regression. 
We quantified overall performance in terms of overall relative root mean squared estimation error (relative RMSE) for the test set.
If $\hat{Y}$ is the estimated LRV and $Y$ is the true LRV, the relative RMSE is ($E_{\rm rel}$) defined as $E_{\rm rel} = \sqrt{\left\langle\left((\hat{Y}-Y)/Y\right)^2\right\rangle}$,
where the expectation is taken over all test stimuli at all LRVs.
The height of each bar gives the relative RMSE for each method and condition.
Note first that AMA improves upon linear regression for all conditions but the first, where performance for both methods is essentially perfect.
This is expected since AMA makes optimal use of the stimulus information under a wider range
of conditions than linear regression.
Moreover, the linear regression method we use as a baseline here
only has access to cone isomerizations at image locations corresponding to the target object, while AMA has access to isomerizations from the entire image patch.
Second, the fact that performance for Condition 2 is very poor tells us that that when there is substantial intensity variation in the illuminant, achieving luminance constancy
directly from the cone isomerizations is difficult.
Performance improves greatly when we consider the same condition but when the cone isomerizations are contrast normalized. 
Indeed, AMA performance for this case is, as in Condition 1, essentially perfect.
Note, however, that Condition 2 does not include variation in the background surface reflectances.
Such reflectances vary scene to scene in natural viewing.
The results for Condition 3 show performance when we add such variation into the training and test sets.
The error here represents the overall level of luminance constancy that our methods achieve for the
most realistic condition that we tested.

@@

\section{Discussion} \label{Discussion}
In this paper, we studied luminance constancy using naturalistic images.
We used a software pipeline, which was developed for this work, to render datasets of multispectral images from scene descriptions.
Because we rendered the images, we were able to labeled each image by the luminance reflectance value (LRV) of a target object
Across images, we varied the LRV of the target object and its relative surface reflectance spectrum, the spectral power distributions of the light sources, and the reflectance
spectra of the background objects in the scene.
This variation was based on statistical models of natural surface reflectance and illumination spectra.
We used the labeled datasets to learn estimators for target object LRV.
We studied the performance of these estimators as we systematically manipulated the spectral variations in the datasets.

ima labeled by the luminance of a target object in each image. VWCC creates 2D multispectral images of 3D scenes based on information provided by the user. Such information include the shape of the 3D space, the shape, size, position and material properties of the objects and light sources in the scene. In our work, we assigned these geometrical and spectral properties using random sampling from available databases of natural objects and light sources. Thus, we generated a large database of labeled hyper-spectral images, with naturalistic scene properties, where we have the ground truth information about every surface and light source. Such databases, of even a few images, are nearly impossible to obtain for natural scenes. This is because it is very hard to measure the surface reflectance and the illumination spectrum at every point in a scene, even for well controlled lab environments.

We used the software generated labeled images in a supervised manner to understand luminance constancy under four types of spectral manipulations (Fig.~\ref{fig:introExampleFigure}-Fig.~\ref{fig:studiedCases}). We found that target LRV can be recovered through simple manipulation of the retinal cone responses for cases where the spectral manipulations are restricted to at most two of the three possible sources of manipulations: the target object surface reflectance, the background object surface reflectances, or the light source illumination spectra. When variations are allowed only in the target object surface reflectance spectrum, or only the target object and the background objects surface reflectance spectra, the luminance can be estimated directly from the response of retinal cone photoreceptors. If the background spectra is fixed, but the target object reflectance spectrum and the light source illumination spectra are varying, the luminance can be obtained from the contrast between the cone response corresponding to the target and the background. When all three spectral variations are allowed simultaneously, one needs additional information to obtain the target LRV.

Summary

Effect of background. multiple reflection

center surround RFs

LM

@@

% NOT SURE WE NEED ANY OF THIS.  PROBABLY NOT HERE IF WE DO
%Even within the special case of luminance constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by Fig.~\ref{fig:introExampleFigure} and the three panels of Fig.~\ref{fig:studiedCases}. These conditions are: 1. variation in target object spectrum with fixed light source and background objects spectra (Fig.~\ref{fig:introExampleFigure}), 2. variation in target object spectrum and background objects spectra, with fixed light source  (Fig.~\ref{fig:backgroundVarying}), 3. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.~\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.~\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \nameref{Methods}). In this paper, we focus on the problem of luminance constancy within scene ensembles where these geometric factors are held fixed.
%
%We show that for conditions that have spectral variation only either in target object surface reflectance or background object surface reflectance or the power spectrum of the light sources, luminance can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (Condition 1) or just the target and background objects reflectance spectra (Condition 2), LRV can be recovered directly from the light reflected off the target object. For Condition 3, where both the target object reflectance and light source spectra vary, LRV can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (Condition 4), the standard luminance can not be recovered simply from the cone response or the contrast. We have used AMA with a Bayesian estimator to estimate standard luminance of the target object. Using this method, we can recover the luminance of the target object with $\sim 15\%$ relative root mean square error.  Our analysis shows that the optimal receptive fields for estimating the luminance of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, luminance estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one condition on the cone response of other conditions shows that the receptive field of the complex conditions perform well on images of simpler conditions. We discuss the implications of such receptive fields on luminance constancy.

>>>>

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. 
But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. 
While this invariance detection ability is behaviorally very significant, its mechanism remains largely unknown. 
To understand the computational principles that lead to invariant color perception, 
here we study the perception of object light reflectance value (LRV) under variations in naturalistic scenes. 
Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the 
perceived LRV of a target object. 
We have developed a software that renders naturalistic multispectral images of 3D scenes.
The software provides precise control over the geometrical and spectral factors that make a scene. 
We label the rendered images with the luminance of a specific object in the scene. 
Next, we simulate the response of the retinal cones to these multispectral images using 
an accurate model of the early visual system. 
We then use supervised learning methods on these labeled cone responses to 
identify the computations that lead to accurate LRV estimation under changes in spectral condition. 
We show that if only either target or target and illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. 
When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, 
a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and  
S cones can recover the luminance within about 13\% RMSE.}

@@

%% Fix the title capitalization in a and b
\subsubsection{Illumination spectra}
% Figure 6
\begin{figure}
\centering
    \begin{subfigure}[b]{0.24 \textwidth}
    \centering
	\caption{Granada dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_a.pdf}
        \label{fig:granadaData}
    \end{subfigure}
	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_b.pdf}
        \label{fig:illuminantSamples}
    \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_c.pdf}
        \label{fig:xyDiagram}
        \end{subfigure}
      	\begin{subfigure}[b]{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure6/Figure6_d.pdf}
        \label{fig:sRGBIlluminant}
    \end{subfigure}
    \caption{{\bf Illumination spectra:} (a) Granada dataset. The spectra are normalized by the mean over wavelength. (b) A representative sample of spectra generated using normalized Granada dataset. (c) CIE xy chromaticity diagram of the normalized dataset and the random samples. (d) sRGB renditions of the samples from Figure~\ref{fig:illuminantSamples}.}
\label{fig:illuminant}
\end{figure}

@@

Because our multivariate Gaussian model is based on the rescaled spectra, it does not embody the large variation in overall intensity of natural daylights.
To obtain illuminant spectra with intensity variation similar to the Granada dataset, we scaled each randomly generated relative spectrum.
The scale factors we sampled at random in the range 0.15 and 150.
Color swatches of samples of illuminant spectra are shown in Figure~\ref{fig:sRGBIlluminant}.

\subsubsection{Surface reflectance spectra}
%% Fix the title capitalization in a and b
% Figure 7 Surface Reflectance Data and Model
\begin{figure}
\centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
            \caption{Munsell and Vrhel dataset}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_a.pdf}
        \label{fig:reflectanceSpectra}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Statistical model}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_b.pdf}
        \label{fig:reflectanceSamples}
    \end{subfigure}
    \begin{subfigure}{0.24 \textwidth}
    \centering
    \caption{CIE xy chromaticity}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_c.pdf}
        \label{fig:xyChroReflectance}
    \end{subfigure}    
    \centering
	\begin{subfigure}{0.24 \textwidth}
    \centering
        \caption{Color swatches}
        \includegraphics[width=\textwidth]{../FiguresDraft4/Figure7/Figure7_d.pdf}
        \label{fig:backgroundSwatches}
    \end{subfigure}
    \caption{{\bf Surface reflectance spectra:} (a) Surface reflectance spectra from the Munsell and Vrhel datasets.(b) Representative samples of object surface reflectances. (c) CIE xy chromaticity diagram of the Munsell and Vrhel datasets (black) and samples of surface reflectances (red). (d) sRGB renditions of the samples of object surface reflectance under CIE D65 daylight.}
\label{fig:surfaceReflectanceGeneration}
\end{figure}

@@

@@

@@

Figure~\ref{fig:barGraphs} summarizes luminance constancy performance across the conditions, for both AMA and linear regression. 
We quantified overall performance in terms of overall relative root mean squared estimation error (relative RMSE) for the test set defined as:
\begin{align}
E_{\rm rel} = \sqrt{\left\langle\left(\frac{\rm{LRV_{Estimated}-LRV_{True}}}{\rm{LRV_{True}}}\right)^2\right\rangle}
\label{eq:relRMSe}
\end{align}
where the expectation is taken over all test stimuli at all LRVs.
The height of each bar gives the relative RMSE for each method and condition.
Note first that AMA improves upon linear regression for all conditions but the first, where performance for both methods is essentially perfect.
This is expected since AMA makes optimal use of the stimulus information under a wider range
of conditions than linear regression.
Moreover, the linear regression method we use as a baseline here
only has access to cone isomerizations at image locations corresponding to the target object, while AMA has access to isomerizations from the entire image patch.
Second, the fact that performance for Condition 2 is very poor tells us that that when there is substantial intensity variation in the illuminant, achieving luminance constancy
directly from the cone isomerizations is difficult.
Performance improves greatly when we consider the same condition but when the cone isomerizations are contrast normalized. 
Indeed, AMA performance for this case is, as in Condition 1, essentially perfect.
Note, however, that Condition 2 does not include variation in the background surface reflectances.
Such reflectances vary scene to scene in natural viewing.
The results for Condition 3 show performance when we add such variation into the training and test sets.
The error here represents the overall level of luminance constancy that our methods achieve for the
most realistic condition that we tested.

@@

\section{Discussion} \label{Discussion}
In this paper, we studied luminance constancy using naturalistic images.
We used a software pipeline, which was developed for this work, to render datasets of multispectral images from scene descriptions.
Because we rendered the images, we were able to labeled each image by the luminance reflectance value (LRV) of a target object.
Across images, we varied the LRV of the target object, its relative surface reflectance spectrum, 
the spectral power distributions of the light sources, and the reflectance spectra of the background objects in the scene.
These variations were based on statistical models of natural surface reflectance and illumination spectra.
We used the labeled datasets to learn estimators for target object LRV.
We studied the performance of these estimators as we systematically manipulated the 
spectral variations in the datasets.
We show that the target object LRV can be recovered almost perfectly if only the target 
object spectrum or the target object spectrum and the illuminant spectra varies.
In the first case, the LRV can be estimated from the cone isomerizations and in the second case 
from the cone isomerization contrast between the target and the background objects.
If, in addition, the background object spectra also vary, the cone contrast no longer 
proportional to the target object LRV.
In this case, the estimators can predict the LRV with relative RMSE ~13\%.

The variability in the final condition results from changes in background object spectra.
Change in background object spectra can affect the estimation in two ways.
First, it will effect the estimation of average background isomerization in the contrast estimation.
Second, it can also effect the total isomerization in cones corresponding to the target object through 
secondary reflections from the background objects.
While the first effect can be minimized by a better sampling of the background, 
the effect of secondary reflections can not be removed through simple manipulations.
This effect has largely been ignored in the field. 
We show that such effects are significant.

A major reason that such effects have been ignored in the study of color constancy is that we have 
lacked tools through which their effects could be quantified.
In this regard, a major contribution of this work has been to develop a computational package that
produces naturalistic images with precise control over all factors that define the scene in the image.
With this software we can systematically vary these factors and precisely quantify their effects on constancy.

@@
